Co-citations in context

### 2019-06-11 cocit_r.tex
bold restructure

% CREATED A BRANCH FOR SAFETY %
%%%%%%%%%%% Article Set-Up %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Article Type:
%% Default is Research.

@@

Our analysis of these semantically-themed citation networks challenges the assertion of universality. We demonstrate reduced model misspecification and observe that while HC remains highly correlated with hit articles in the immunology and metabolism datasets, HN is not. In addition, while HN is highly correlated with hit articles in applied physics, HC is not. Thus, disciplinary networks are different from each other and the full WoS network. Beyond variation across disciplines and the case for contextual analysis, we found that the categories demonstrating the highest percentage of hits are not robust with respect to varying parameters that define hit articles and the threshold for highly novel citation patterns.

\emph{these remaining sentences require some work}
The varying citation patterns when sampling from a disciplinary-focused dataset versus a broader dataset due, was a cause of journal pair z-scores changing signs in 28.6\% of instances: the effect of any one of these journal pairs on an article's novelty or conventionality is contradictory between broad and, more narrow disciplinary datasets. We contend that the interpretation due to a disciplinary dataset, which preserved citation patterns, is more appropriate and an improvement on current methods.

@@

\subsection{Model Misspecification and the Attributes of Disciplinary Context} \emph{This is where we want to add the additional data about expected values ~2.5\% for AP as an example.} We invoke model mis-specification due to considering disciplinarily irrelevant references as a concern when disciplinary subsets are analyzed using a random graph model that generates expected co-citation frequencies from the WoS superset. A symptom of model misspecification is that random sampling of citations associated with an article do not match with the distribution of citations across the disciplines observed in articles from that discipline. To address this issue, we first assembled three exemplar datasets using keyword searches for the terms `applied physics', `immunology', and `metabolism' (Materials and Methods). Using the WoS extended classificatation of 153 subject areas, we then analyzed the disciplinary composition of the references in these three datasets into the five research areas in the Web of Science (life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities) for the years 1985, 1995, and 2005. We observed that 84.2\% of the references in the applied physics dataset were classified as belonging to the WoS physical sciences research area. Similarly, 93.7\% and 92.1\% of the references in the immunology and metabolism datasets belonged to the life sciences and biomedicine research area, indicating strongly that references are more likely belong to the same research area as the publications that cite them. Our observations are in agreement with \cite{wallace_lariviere_gingras_2012} who found that, often, a majority of an article's citations are from the specialty of the article, although that percentage varied among disciplines in the eight specialties they investigated: from approximately 39\% to 89\% for 2006.

@@

Further, we stratified publications in these datasets by percentiles using the number of citations accumulated in the first 8 years since publication and according to conventionality and novelty in a second dimension Fig \ref{fig:Fig1} and note that [INSERT WHATEVER WE NOTE].

\subsection{Novelty and Conventionality as Determinants of Impact in Disciplinary Contexts}

\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[width=1.5in]{Fig2H01N10.pdf} & \includegraphics[width=1.5in]{Fig2H10N10.pdf} \\
(a) & (b) \\
\end{tabular}
\caption{Effect of Context on Journal-Pair z-scores and Categorical Hit Rates: Immunology, Applied Physics, and WoS for 1995. Panels (a) and (b) show hit rates for the LNLC, LNHC, HNLC, and HNHC categories for the immunology, applied physics, and WoS datasets when hit articles are defined as the top 1\% and top 10\% of articles, respectively.  Novelty in both panels is defined at the 10th percentile of articles' z-score distributions. The number of data points in the immunology, applied physics, and WoS data sets are 21,917, 18,305, and 476,288, respectively.  The results for the WoS data set mirrored previous results from Uzzi et al. ~\cite{uzzi_atypical_2013} where the highest hit rate was for the HNHC category.  The highest hit rates for the immunology dataset, in contrast, are in the LNHC category. The LNHC category often had the highest hit rate for the immunology dataset for various parameter settings with the HCHN category having the highest hit rate in other cases: metabolism results are similar. The applied physics dataset shows further contrast as the hit rate for it was highest in the HNLC category.  The HNLC category demonstrated the highest hit rate often for applied physics: otherwise the highest hit rate was for HNHC articles.}
\label{fig:Fig2}
\end{figure*}

@@

We conclude thathigh hit rates in the HNHC category does not hold generally for all datasets and that conventionality flavored with novelty is not always indicative of impactful research. The z-scores that change sign across datasets is either contradictory or an acceptable variation due to the difference in reference sets. We contend that it is the former because the inappropriate substitution of citations with those from disciplines that are implausible and because the observed citation frequencies are ignored. That fewer z-scores are negative in the WoS dataset relative to the immunology dataset may be directly due to the uniform sampling of references whereby many resulting journal pairs are never observed in the literature and so that the expected frequencies of those observed pairs have lower expected values, thus biasing their z-scores upward. In addition, treating citation as a set versus a multiset means that a random model will sample popular citations downward, thus further increasing their z-scores.\emph{need confirmation from Jim here}.

@@

\newpage
%%%%%%%%%%%%%%%%%%%%%%%
%% The bibliography

>>>>

%%%%%%%%%%% Article Set-Up %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Article Type:
%% Default is Research.

@@

Our analysis of these semantically-themed citation networks challenges the assertion of universality. We demonstrate reduced model misspecification and observe that while HC remains highly correlated with hit articles in the immunology and metabolism datasets, HN is not. \emph{needs another sentence referencing Fig 2}. In addition, while HN is highly correlated with hit articles in applied physics, HC is not. Thus, disciplinary networks are different from each other and the full WoS network. Beyond variation across disciplines and the case for contextual analysis, we found that the categories demonstrating the highest percentage of hits are not robust with respect to varying parameters that define hit articles and the threshold for highly novel citation patterns.

@@

\subsection{Model Misspecification and the Attributes of Disciplinary Context} \emph{This is where we want to add the additional data about expected values ~2.5\% for AP as an example.} We invoke model mis-specification due to considering disciplinarily irrelevant references as a concern when disciplinary subsets are analyzed using a random graph model that generates expected co-citation frequencies from the WoS superset. A symptom of model misspecification is that random sampling of citations associated with an article do not match with the distribution of citations across the disciplines observed in articles from that discipline. To address this issue, we first assembled three exemplar datasets using keyword searches for the terms `applied physics', `immunology', and `metabolism' (Materials and Methods). Using the WoS extended classificatation of 153 subject areas, we then analyzed the disciplinary composition of the references in these three datasets into the five research areas in the Web of Science (life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities) for the years 1985, 1995, and 2005. We observed that 84.2\% of the references in the applied physics dataset were classified as belonging to the WoS physical sciences research area, which is comparable to 79.5\% for the references in the physical sciences area. Similarly, 93.7\% and 92.1\% of the references in the immunology and metabolism datasets belonged to the life sciences and biomedicine research area, again comparable to 89.9\% for the references in the life sciences and biomedicine area. \emph{Add in Sitaram's new data}. These data indicate strongly that references are more likely belong to the same research area as the publications that cite them. Our observations are in agreement with \cite{wallace_lariviere_gingras_2012} who found that, often, a majority of an article's citations are from the specialty of the article, although that percentage varied among disciplines in the eight specialties they investigated: from approximately 39\% to 89\% for 2006.

@@

\subsection{Novelty and Conventionality as Determinants of Impact in Disciplinary Contexts}
\begin{figure*}
\centering
\begin{tabular}{ccc}
\includegraphics[width=1.5in]{Fig2H01N10.pdf} & \includegraphics[width=1.5in]{Fig2H10N10.pdf} \\
(a) & (b) \\
\end{tabular}
\caption{Effect of Context on Journal-Pair z-scores and Categorical Hit Rates: Immunology, Applied Physics, and WoS for 1995. Panels (a) and (b) show hit rates for the LNLC, LNHC, HNLC, and HNHC categories for the immunology, applied physics, and WoS datasets when hit articles are defined as the top 1\% and top 10\% of articles, respectively.  Novelty in both panels is defined at the 10th percentile of articles' z-score distributions. The number of data points in the immunology, applied physics, and WoS data sets are 21,917, 18,305, and 476,288, respectively.  The results for the WoS data set mirrored previous results from Uzzi et al. ~\cite{uzzi_atypical_2013} where the highest hit rate was for the HNHC category.  The highest hit rates for the immunology dataset, in contrast, are in the LNHC category. The LNHC category often had the highest hit rate for the immunology dataset for various parameter settings with the HCHN category having the highest hit rate in other cases: metabolism results are similar. The applied physics dataset shows further contrast as the hit rate for it was highest in the HNLC category.  The HNLC category demonstrated the highest hit rate often for applied physics: otherwise the highest hit rate was for HNHC articles.}
\label{fig:Fig2}
\end{figure*}

@@

\emph{needs to be expanded}
We conclude that high hit rates in the HNHC category do not hold generally for all datasets and that conventionality flavored with novelty is not always indicative of impactful research. The z-scores that change sign across datasets is either contradictory or an acceptable variation due to the difference in reference sets. We contend that it is the former because the inappropriate substitution of citations with those from disciplines that are implausible and because the observed citation frequencies are ignored. That fewer z-scores are negative in the WoS dataset relative to the immunology dataset may be directly due to the uniform sampling of references whereby many resulting journal pairs are never observed in the literature and so that the expected frequencies of those observed pairs have lower expected values, thus biasing their z-scores upward. In addition, treating citation as a set versus a multiset means that a random model will sample popular citations downward, thus further increasing their z-scores.\emph{need confirmation from Jim here}.

@@

\newpage

\section{Residual}
\emph{these remaining sentences require some work}
The varying citation patterns when sampling from a disciplinary-focused dataset versus a broader dataset due, was a cause of journal pair z-scores changing signs in 28.6\% of instances: the effect of any one of these journal pairs on an article's novelty or conventionality is contradictory between broad and, more narrow disciplinary datasets. We contend that the interpretation due to a disciplinary dataset, which preserved citation patterns, is more appropriate and an improvement on current methods.

%%%%%%%%%%%%%%%%%%%%%%%
%% The bibliography
